---
title: "powerLMM"
output: html_document
date: "2024-02-08"
---

To conduct the power sensitivity test in R, we first generated a simulated dataset using the mean and SD from VPEEG study. The sample size of dataset was based on previous studies. Then, we created linear mixed model (LMM) following our analysis plan. Then, we adjusted the estimate of each predictor manually (e.g., fixef(model1)["condition"] <- 0.7) and calculated the corresponding effect size r. After adjusting the estimate, we ran power analysis using “powerSim” and then checked if the power was higher than 80%. By setting different estimates/effect sizes, we were able to learn the sample size can detect what effect size with 80%. If the power is lower than 80%, we would increase the estimate/effect size. If the power is higher than 80%, we would decrease the estimate/effect size. 

```{r data simulation}
library(lmerTest)

# Set seed for reproducibility
set.seed(42)

# Number of subjects
n_subjects <- 80 # to examine how large effect size we can detect with this sample size

# Generate random values for diversity (one for each subject)
diversity <- rnorm(n_subjects, mean = 0.24, sd = 0.27) # based on VPEEG network diversity

# Number of observations per subject
n_obs_per_subject <- 4 # 2 windows for each speaker, 4 in total

# Create a pattern of condition
condition <- rep(c(1, 0), times = n_subjects * n_obs_per_subject / 2) # 1 indicates own-race, 0 indicates other-race

# Repeat diversity values for each observation within the subject
diversity <- rep(diversity, each = n_obs_per_subject) # each subject should have the same diversity value

# Create a data frame with generated data
data <- data.frame(diversity = diversity, condition = condition)

# Generate random values for the grouping variable (ID)
data$ID <- factor(rep(1:n_subjects, each = n_obs_per_subject))

data$theta <- rnorm(n_subjects * n_obs_per_subject, 0.08, 1.81) # generate scores based on the VPEEG
data$mu <- rnorm(n_subjects * n_obs_per_subject, -0.53, 1.26) # generate scores based on the VPEEG
data$FAA <- rnorm(n_subjects * n_obs_per_subject, .003, .03)  # generate scores based on the VPEEG

# Generate child language background
data$child_language <- rep(sample(c(0, 1), n_subjects, replace = TRUE), each = n_obs_per_subject)

# Inspect the first few rows
head(data)

```


```{r mu language condition effect}
library(lmerTest)
library(simr)
# Fit the linear mixed-effects model
model1 <- lmer(mu ~ 1 + condition + diversity + 
                diversity:condition + 
                child_language + 
                child_language:condition + child_language:condition:diversity+
                (1 | ID), 
              data = data)

# Summary of the model
summary(model1)

fixef(model1)["condition"]
fixef(model1)["condition"] <- 0.83 # test the effect size to reach 80% power
isSingular(model1)
powerSim(model1, test=fixed("condition"),nsim=1000)

fixef(model1)
all_t<-summary(model1)$coefficients[, "t value"]
all_df<-summary(model1)$coefficients[, "df"]
t <- all_t[2]
df <- all_df[2]
r <- sqrt(t^2/(t^2+df))
r
```

```{r mu diversity and language interaction effect}
fixef(model1)["condition:diversity"]
fixef(model1)["condition:diversity"] <- -2.1 # test the effect size to reach 80% power

powerSim(model1, test=fixed("condition:diversity"))

fixef(model1)
all_t<-summary(model1)$coefficients[, "t value"]
all_df<-summary(model1)$coefficients[, "df"]
t <- all_t[5]
df <- all_df[5]
r <- sqrt(t^2/(t^2+df))
r
```
```{r mu child language background and language interaction effect}
fixef(model1)["condition:child_language"]
fixef(model1)["condition:child_language"] <- -1.0 # test the effect size to reach 80% power

powerSim(model1, test=fixed("condition:child_language"))

fixef(model1)
all_t<-summary(model1)$coefficients[, "t value"]
all_df<-summary(model1)$coefficients[, "df"]
t <- all_t[6]
df <- all_df[6]
r <- sqrt(t^2/(t^2+df))
r
```

```{r FAA language condition effect}
library(lmerTest)
library(simr)
# Fit the linear mixed-effects model
model1 <- lmer(FAA ~ 1 + condition + diversity + 
                diversity:condition + 
                child_language + 
                child_language:condition + child_language:condition:diversity+
                (1 | ID), 
              data = data)

# Summary of the model
summary(model1)

fixef(model1)["condition"]
fixef(model1)["condition"] <- 0.02 # test the effect size to reach 80% power

powerSim(model1, test=fixed("condition"))

fixef(model1)
all_t<-summary(model1)$coefficients[, "t value"]
all_df<-summary(model1)$coefficients[, "df"]
t <- all_t[2]
df <- all_df[2]
r <- sqrt(t^2/(t^2+df))
r
```
```{r FAA diversity and language interaction effect}
fixef(model1)["condition:diversity"]
fixef(model1)["condition:diversity"] <- -0.055 # test the effect size to reach 80% power

powerSim(model1, test=fixed("condition:diversity"))

fixef(model1)
all_t<-summary(model1)$coefficients[, "t value"]
all_df<-summary(model1)$coefficients[, "df"]
t <- all_t[5]
df <- all_df[5]
r <- sqrt(t^2/(t^2+df))
r
```
```{r FAA child language background and language interaction effect}
fixef(model1)["condition:child_language"]
fixef(model1)["condition:child_language"] <- -0.025 # test the effect size to reach 80% power

powerSim(model1, test=fixed("condition:child_language"))

fixef(model1)
all_t<-summary(model1)$coefficients[, "t value"]
all_df<-summary(model1)$coefficients[, "df"]
t <- all_t[6]
df <- all_df[6]
r <- sqrt(t^2/(t^2+df))
r
```
